Abstract Binding Trees

### 2015-10-21 paper.tex
Fix math font
Fix forgotten rule
Change notation for valences and bound terms
Add todos
correct the size of some delimiters; not sure if this is an improvement
Try and resolve #6

% fonts
% \setromanfont[Mapping=tex-text, Ligatures=Rare, Numbers=OldStyle]{Linux Libertine O}
\setmathfont{Asana Math}

@@

\newcommand\MkValence[3]{#1;#2\,.\, #3}
\newcommand\MkBTm[3]{#1;#2\,.\, #3}
\newcommand\MkArity[2]{\left(#1\right) #2}
\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}

@@

Let $\Symbols$ be an infinite set of symbols; let $\Finsets$ be the free
cocartesian category over $\Symbols$. Then, fix a covariant presheaf of operators
$\Operators:\Finsets\times\Arities\to\Sets$ such that the arrows in $\Finsets$
lift to renamings of operators' parameters; via the Grothendieck construction,
we can also consider the set $\int\Operators$ of operators
$(U,\varrho,\vartheta)$ for $\Member{\vartheta}{\Operators(U,\varrho)}$.

\[
  \infer{
    \IsOperator{U}{\vartheta}{\varrho}
  }{
    \Member{\vartheta}{\Operators(U,\varrho)}
  }
\]

The judgment $\IsOperator{U}{\vartheta}{\varrho}$ enjoys the structural
properties of weakening and exchange via the functoriality of
$\Operators$.

@@

Operators are defined by specifying the fibres of $\int\Operators$ in which
they reside. For instance, consider the lambda calculus with a single sort,
$\SortExp$; we give its signature by asserting the following about its
operators:
\begin{gather*}
  \IsOperator{U}{\lambda}{
    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp}
  }\\
  \IsOperator{U}{\mathsf{ap}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{\SortExp}
  }
\end{gather*}

@@

\begin{gather*}
  \IsOperator{U}{\mathsf{decl}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\SortExp}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }\\
  \IsOperator{U,u}{\mathsf{get}[u]}{
    \MkArity{
      \cdot
    }{
      \SortExp
    }
  }\\
  \IsOperator{U,u}{\mathsf{set}[u]}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }
\end{gather*}

@@

\begin{gather*}
  \infer[\textit{var}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{s}
  }{
    \Lookup{\Gamma}{x}{s}
  }\\[6pt]
  %
  \infer[\textit{mvar}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
    }{s}
  }{
    \begin{array}{l}
      \Lookup{\Omega}{\MV{m}}{\MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}}\\
      \HypJ{\Lookup{\Upsilon}{u_i}{p_i}}{\Leq{i}{m}}\\
      \HypJ{
        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }\\[6pt]
  %
  \infer[\textit{app}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \App{\vartheta}{E_0,\dots,E_n}
    }{s}
  }{
    \begin{array}{l}
      \IsOperator{\Dom{\Upsilon}}{\vartheta}{
        v_1,\dots,v_n
      }\\
      \HypJ{
        \IsBTm{\Omega}{\Upsilon}{\Gamma}{E_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }\\[6pt]
  \infer[\textit{abs}]{
    \IsBTm{\Omega}{\Upsilon}{\Gamma}{
      (\MkBTm{\vec{u}}{\vec{x}}{E})
    }{
      (\MkValence{\vec{p}}{\vec{q}}{s})
    }
  }{
    \IsAbt{\Omega}{\Upsilon, \vec{u}:\vec{p}}{\Gamma,\vec{x}:{\vec{q}}}{
      E
    }{
      s
    }
  }
\end{gather*}

>>>>

% fonts
% \setromanfont[Mapping=tex-text, Ligatures=Rare, Numbers=OldStyle]{Linux Libertine O}
\setmathfont{Asana-Math.otf}

@@

\newcommand\MkValence[3]{\left\{#1\right\}\left[#2\right]\,.\, #3}
\newcommand\MkBTm[3]{\left\{#1\right\}\left[#2\right]\,.\, #3}
\newcommand\MkArity[2]{\left(#1\right) #2}
\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}

@@

Let $\Symbols$ be an infinite set of symbols; let $\Finsets$ be the free
cocartesian category over $\Symbols\times\Sorts$. Then, fix a covariant presheaf of operators
$\Operators:\Finsets\times\Arities\to\Sets$ such that the arrows in $\Finsets$
lift to renamings of operators' parameters; via the Grothendieck construction,
we can also consider the set $\int\Operators$ of operators
$(U,\varrho,\vartheta)$ for $\Member{\vartheta}{\Operators(\Upsilon,\varrho)}$.

\[
  \infer{
    \IsOperator{\Upsilon}{\vartheta}{\varrho}
  }{
    \Member{\vartheta}{\Operators(\Upsilon,\varrho)}
  }
\]

The judgment $\IsOperator{\Upsilon}{\vartheta}{\varrho}$ enjoys the structural
properties of weakening and exchange via the functoriality of
$\Operators$.

@@

Operators are defined by specifying the fibres of $\int\Operators$ in which
they reside. For instance, consider the lambda calculus with a single sort,
$\SortExp$; we give its signature by asserting the following about its
operators:
\begin{gather*}
  \IsOperator{\Upsilon}{\lambda}{
    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp}
  }\\
  \IsOperator{\Upsilon}{\mathsf{ap}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{\SortExp}
  }
\end{gather*}

@@

\begin{gather*}
  \IsOperator{\Upsilon}{\mathsf{decl}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\SortExp}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }\\
  \IsOperator{\Upsilon,u:\SortExp}{\mathsf{get}[u]}{
    \MkArity{
      \cdot
    }{
      \SortExp
    }
  }\\
  \IsOperator{\Upsilon,u:\SortExp}{\mathsf{set}[u]}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{
      \SortExp
    }
  }
\end{gather*}

@@

\begin{gather*}
  \infer[\textit{var}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{x}{s}
  }{
    \Lookup{\Gamma}{x}{s}
  }\\[6pt]
  %
  \infer[\textit{mvar}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \MApp{\MV{m}}{u_0,\dots,u_m}{M_0,\dots,M_n}
    }{s}
  }{
    \begin{array}{l}
      \Lookup{\Omega}{\MV{m}}{\MkValence{p_0,\dots,p_m}{q_0,\dots,q_n}{s}}\\
      \HypJ{\Lookup{\Upsilon}{u_i}{p_i}}{\Leq{i}{m}}\\
      \HypJ{
        \IsAbt{\Omega}{\Upsilon}{\Gamma}{M_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }\\[6pt]
  %
  \infer[\textit{app}]{
    \IsAbt{\Omega}{\Upsilon}{\Gamma}{
      \App{\vartheta}{E_0,\dots,E_n}
    }{s}
  }{
    \begin{array}{l}
      \IsOperator{\Upsilon}{\vartheta}{
        v_1,\dots,v_n
      }\\
      \HypJ{
        \IsBTm{\Omega}{\Upsilon}{\Gamma}{E_i}{q_i}
      }{
        \Leq{i}{n}
      }
    \end{array}
  }\\[6pt]
  \infer[\textit{abs}]{
    \IsBTm{\Omega}{\Upsilon}{\Gamma}{
      (\MkBTm{\vec{u}}{\vec{x}}{E})
    }{
      (\MkValence{\vec{p}}{\vec{q}}{s})
    }
  }{
    \IsAbt{\Omega}{\Upsilon, \vec{u}:\vec{p}}{\Gamma,\vec{x}:{\vec{q}}}{
      E
    }{
      s
    }
  }
\end{gather*}
